---
title: "sqlcaseR: Building long CASE WHEN statements for SQL interfaces in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sqlcaseR: Building long CASE WHEN statements for SQL interfaces in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
**Version 0.1.0**

***Leoson Hoay <br>16 Dec 2022***


## Introduction
This module was born out of my genuine frustration while constructing an
extremely long CASE WHEN...THEN statement to re-label categorical variables.
It is most helpful for folks who intend to work with SQL directly in the R
environment, likely with a SQL connector such as RODBC or RSQLite. 

Instead of manually inputting WHENs and THENs to replace/map values, this
library does it for you if you provide it with a mapping CSV file that contains
the original values in the first column, and the values to map to in the second
column. This version only supports CSV files for now, but support for other file
formats is planned.

<br>
Go from this:
                  
| | |
|--------------------------------|-------------------------------|
| Hotel/Motel	                   | Living in Shelter/Hotel/Motel |
| Homeless Shelter	             | Living in Shelter/Hotel/Motel |
| Homeless Status Not Applicable | Not Homeless                  |
| N/A	                           | Not Homeless                  |
| No	                           | Not Homeless                  |
| Homeless, Doubled-Up	         | Doubled Up                    |

To this:

```{}
CASE WHEN 'Hotel/Motel' THEN 'Living in Shelter/Hotel/Motel'
 WHEN 'Homeless Shelter' THEN 'Living in Shelter/Hotel/Motel'
 WHEN 'Homeless Status Not Applicable' THEN 'Not Homeless'
 WHEN 'N/A' THEN 'Not Homeless'
 WHEN 'No' THEN 'Not Homeless'
 WHEN 'Homeless, Doubled-Up' THEN 'Doubled Up'
```

## Demonstration

```{r setup}
library(sqlcaser)
```

The package assumes that the user has a mapping CSV file similar to the example
below: 

```{r}
samp <- system.file("extdata", "sample.csv", package = "sqlcaser")
mapping <- read.csv(samp)
mapping
```

The function **casewhen()** takes the ***file path*** of the mapping file as input, 
and returns the CASE statement as a string, while printing it to the console
as well.

```{r}
statement <- casewhen(samp)
```

The user can then easily include it as part of the SQL query:

```{r}
query <- paste("SELECT id, ", statement, " END AS status "," \nFROM table;")
cat(query)
```

## Sample Data
A sample mapping file is provided in this package. The file path can be accessed
as follows:

```{r}
samplepath <- system.file("extdata", "sample.csv", package = "sqlcaser")
```

## Functions

***casewhen(inputfile)***

**description**

This function constructs a CASE WHEN statement from a mapping CSV file.
It assumes that the first column of the mapping file contains the original
CASE WHEN values, and the second column contains the THEN values (the values to
be mapped to.)

**Usage**

casewhen(inputfile)

**Arguments**

*inputfile* Path to the mapping file

**Value**

A string that represents the constructed CASE WHEN statement


## Installation

## Acknowledgments
Much thanks to a couple of my colleagues at Learning Collider, Nitya Raviprakash
and Jasmin Dial, who provided healthy discussion around my misery of
constructing long SQL queries. Credit is also due to Kevin Flerlage, whose
efforts in automating this process in Excel should be commended and partially
inspired this package.

## Citation and License
If desired, cite the package using:

```{}
citation("sqlcaser")
```

License: MIT License